Ceph Parallel File System Evaluation Report

### 2013-09-20 abstract.tex
Brad's attempt to cut to 5 pages
So here is most of my attempt to cut this paper to 5 pages.
I amde it to about 5.7 pages.  I could go after it some more,
but this is a good first draft point to check and see if we need
to add back anything I took out.

I also added the IEEEabbrv.bib file since I couldn't build
without it.

Ceph is an open-source and emerging parallel distributed file and storage
system technology. By design, Ceph assumes running on unreliable and commodity
storage and network hardware and provides reliability and fault-tolerance
through controlled object placement and data replication.

We evaluated the Ceph
technology for scientific high-performance computing (HPC) environments.  This paper
presents our evaluation methodology, experiments, results and observations from
mostly parallel I/O performance and scalability perspectives.  Our work made
two unique contributions. First, our evaluation is performed under a realistic
setup for a large-scale capability HPC environment using a commercial
high-end storage system.
% (Data Direct Network's SFA10K).
%although typical in large-scale HPC environment, but atypical for average
%commercial off-the-shelf setup.  
Second, our path of investigation, tuning efforts, and findings made direct
contributions to Ceph's development and improved code quality, scalability, and
performance. These changes should also benefit both Ceph and HPC communities at
large. Throughout the evaluation, we observed that Ceph still is an evolving
technology under fast-paced development and showing great promises.

>>>>

Ceph is an emerging open-source parallel distributed file and storage
system. By design, Ceph is designed to leverage unreliable and commodity
storage and network hardware, and provides reliability and fault-tolerance
via controlled object placement and data replication.  Here we evaluate Ceph
for scientific high-performance computing (HPC) environments.  This paper
presents our evaluation methodology, experiments, results and observations from
mostly parallel I/O performance and scalability perspectives.  Our work makes
two unique contributions. First, our evaluation is performed under a realistic
setup for a large-scale capability HPC environment using a commercial
high-end storage system.
% (Data Direct Network's SFA10K).
%although typical in large-scale HPC environment, but atypical for average
%commercial off-the-shelf setup.  
Second, our path of investigation, tuning efforts, and findings made direct
contributions to Ceph's development and improved code quality, scalability, and
performance. These changes should benefit both Ceph and the HPC community
at large. Throughout the evaluation, we observe that Ceph is an evolving
technology under fast-paced development and shows great promises.

### 2013-09-20 baseline.tex
Brad's attempt to cut to 5 pages
So here is most of my attempt to cut this paper to 5 pages.
I amde it to about 5.7 pages.  I could go after it some more,
but this is a good first draft point to check and see if we need
to add back anything I took out.

I also added the IEEEabbrv.bib file since I couldn't build
without it.

\section{Establishing A Baseline Performance}
\label{sec:baseline}

@@

The lowest layer in the system are the exposed LUNs to the server hosts from
the RAID controllers.  These LUNs are configured as a RAID 6 8+2 array (8 data
disks and 2 \textit{P} and \textit{Q} disks). In our earlier tests we observed
that a 7.2K RPM SAS disk can perform at \textasciitilde140 MB/s for 128 kB
sequential I/O requests and a 7.2K RPM SATA disk can do \textasciitilde36 MB/s.
For these tests, all disk caches were turned off. Therefore, in 8 data disk
RAID group with 1 MB sequential I/O requests (each disks sees 128 kB chunks of
the request) the expected aggregate performance would be \textasciitilde1.12
GB/s for a SAS RAID group and \textasciitilde288 MB/s for a SATA RAID group, if
there are no caches along the I/O path.  With the caches turned on, especially
the write-back cache on the DDN RAID controller, we observe a significant
performance improvement. In our day-to-day HPC operations we run our storage
systems with write-back cache turned on on RAID controllers, if and only if,
there is a method of cache-mirroring between the two active-active RAID
controllers. With the write-back cache on we observe \textasciitilde1.4 GB/s
per RAID 6 8+2 SAS group and  \textasciitilde955 MB/s per RAID 6 8+2 SATA
group, for 1 MB sequential I/O.  These two performance numbers will be used as
a baseline in our study and they are presented in
Table~\ref{tbl:block-io-baseline}.  For these tests, the 8+2 RAID 6 groups are
exported to the server hosts using SCSI RDMA Protocol (SRP) over IB protocol
(which is the default for our normal HPC configurations). For a SATA disk group
this translates into roughly 120 MB/s/disk and for a SAS disk group it is 175
MB/s/disk. As can be seen, SATA RAID groups benefit from caching more than SAS
groups.

For above tests the RAID groups are exported over physical 4x IB QDR links.
These links can provide up to 40 Gbits/s signalling rate and 32 Gbits/s actual
data rate (reduction is due to 8/10 encoding scheme). Therefore, each IB QDR
link can perform at 3.2 GB/s, while in practice we see around 3 GB/s as the
best case.

As mentioned earlier, we had 280 SATA disks and 200 SAS disks. Organizing these
into RAID 6 8+2 groups will yield 28 SATA and 20 SAS RAID 6 8+2 groups. We had
4 server hosts in our testbed and evenly distributing these on to 4 server
hosts translates into 7 SATA LUNs and 4 SAS LUNs per server host.

Next, we established the baseline performance for a single server host. We
exercised all seven SATA LUNs concurrently form a single server host at the
block-level. This resulted in an aggregate performance of roughly 2.6 GB/s.
Repeating the same test using only four SAS LUNs yielded around 2.8 GB/s
aggregate performance for 1 MB sequential I/O. These two numbers will be used
as single server host aggregate baseline performance for this study. These are
also presented in Table~\ref{tbl:block-io-baseline}.

It is worth mentioning that we used our in-house developed synthetic benchmark,
\textit{fair-lio} for all our block-level testing.  fair-lio has better
sequential I/O characteristics compared to commonly used XDD benchmark and it
is asynchronous and based on the Linux \verb!libaio! library.

@@

When exercising all 28 SATA LUNs from all four server hosts in parallel, we
observed an 11 GB/s aggregate performance. Same level of performance was
observed using all 20 SAS LUNs from four server hosts in parallel. Comparing
these to the advertised peak performance of 12 GB/s of the DDN SFA10K, we
concluded that our test setup is well configured to drive the backend disk
system at close to full performance and we are limited by the RAID controller
performance. Going forward in this study, 11 GB/s will be used the peak
baseline for the entire test configuration.

@@

Unfortunately there was not enough time to do more detailed analysis of the
network performance. However, as we observed later, RADOS is performing no
more than 8 GB/s driven by four server hosts. This confirms that we have
provisioned enough network bandwidth. In other words, IP over IB is not a
bottleneck in this case.

>>>>

\section{Establishing A Performance Baseline}
\label{sec:baseline}

@@

The lowest layer in the system are the exposed LUNs to the server hosts from
the RAID controllers.  These LUNs are configured as a RAID 6 8+2 array (8 data
disks and 2 \textit{P} and \textit{Q} disks). In prior tests we observed
that a 7.2K RPM SAS disk can perform at \textasciitilde140 MB/s for 128 kB
sequential I/O requests and a 7.2K RPM SATA disk can do \textasciitilde36 MB/s.
For these tests, all disk caches were turned off. Therefore, in 8 data disk
RAID group with 1 MB sequential I/O requests (each disks sees 128 kB chunks of
the request) the expected aggregate performance would be \textasciitilde1.12
GB/s for a SAS RAID group and \textasciitilde288 MB/s for a SATA RAID group, if
there are no caches along the I/O path.  With the caches turned on, especially
the write-back cache on the DDN RAID controller, we observe a significant
performance improvement. In our day-to-day HPC operations we run our storage
systems with write-back cache turned on on RAID controllers, if and only if,
there is a method of cache-mirroring between the two active-active RAID
controllers. With the write-back cache on we observe \textasciitilde1.4 GB/s
per RAID 6 8+2 SAS group and  \textasciitilde955 MB/s per RAID 6 8+2 SATA
group, for 1 MB sequential I/O.  These two performance numbers will be used as
a baseline in our study and they are presented in
Table~\ref{tbl:block-io-baseline}.  For these tests, the 8+2 RAID 6 groups are
exported to the server hosts using SCSI RDMA Protocol (SRP) over IB protocol
(which is the default for our normal HPC configurations). For a SATA disk group
this translates into roughly 120 MB/s/disk and for a SAS disk group it is 175
MB/s/disk. As can be seen, SATA RAID groups benefit from caching more than SAS
groups.

%For above tests the RAID groups are exported over physical 4x IB QDR links.
%These links can provide up to 40 Gbits/s signalling rate and 32 Gbits/s actual
%data rate (reduction is due to 8/10 encoding scheme). Therefore, each IB QDR
%link can perform at 3.2 GB/s, while in practice we see around 3 GB/s as the
%best case.

%As mentioned earlier, we had 280 SATA disks and 200 SAS disks. Organizing these
%into RAID 6 8+2 groups yields 28 SATA and 20 SAS RAID 6 8+2 groups. We had
%4 server hosts in our testbed and evenly distributing these on to 4 server
%hosts translates into 7 SATA LUNs and 4 SAS LUNs per server host.

%Next, we established the baseline performance for a single server host. We
%exercised all seven SATA LUNs concurrently form a single server host at the
%block-level. This resulted in an aggregate performance of roughly 2.6 GB/s.
%Repeating the same test using only four SAS LUNs yielded around 2.8 GB/s
%aggregate performance for 1 MB sequential I/O. These two numbers will be used
%as single server host aggregate baseline performance for this study. These are
%also presented in Table~\ref{tbl:block-io-baseline}.

It is worth mentioning that we used our in-house developed synthetic benchmark,
\textit{fair-lio} for all our block-level testing.  Fair-lio has better
sequential I/O characteristics compared to commonly used XDD benchmark and it
is asynchronous and based on the Linux \verb!libaio! library.

@@

When exercising all 28 SATA LUNs from all four server hosts in parallel, we
observed an 11 GB/s aggregate performance. The same level of performance was
observed using all 20 SAS LUNs from four server hosts in parallel. Comparing
these to the advertised peak performance of 12 GB/s of the DDN SFA10K, we
concluded that our test setup is well configured to drive the backend disk
system at close to full performance and we are limited by the RAID controller
performance. Going forward in this study, 11 GB/s will be used the peak
baseline for the entire test configuration.

@@

%Unfortunately there was not enough time to do more detailed analysis of the
%network performance. However, as we observed later, RADOS is performing no
%more than 8 GB/s driven by four server hosts. This confirms that we have
%provisioned enough network bandwidth. In other words, IP over IB is not a
%bottleneck in this case.

### 2013-09-20 ceph-initial.tex
Brad's attempt to cut to 5 pages
So here is most of my attempt to cut this paper to 5 pages.
I amde it to about 5.7 pages.  I could go after it some more,
but this is a good first draft point to check and see if we need
to add back anything I took out.

I also added the IEEEabbrv.bib file since I couldn't build
without it.

RADOS is a reliable distributed object store, the foundational component for
CephFS file system. There are two types of scaling tests we are interested at
the RADOS layer:

\begin{itemize}
  \item scaling on the number of OSD servers
  \item scaling on the number of OSDs per OSD server
\end{itemize}

Our system setup poses some limitations on the scalability tests we wanted to
perform. In particular, we currently had four OSD servers, eight clients, and
eleven OSD servers per client. The scaling tests, therefore, will be within
these constraints.

@@

During our early tests, a trend that previously had been seen became more
apparent.  During reads, there were periods of high performance followed by
periods of low performance or outright stalls that could last for up to 20
seconds at a time.  RADOS bench recorded poor read performance across
different backend file systems, as shown in
Figure~\ref{fig:rados-tcp-autotune}.

After running diagnostics, we observed that outstanding operations on the
clients were not being shown as outstanding on the OSDs.  This appeared to be
very similar to a problem Jim Schutt at Sandia National labs encountered with
TCP autotuning in the Linux
kernel\cite{Jim:tcptune}.
TCP auto tuning enables TCP window scaling by default and automatically adjusts
the TCP receive window for each connection based on link conditions such as
bandwidth delay product.

@@

Luckily, the fix was fairly straight forward by issuing the following command
on all nodes:

@@

With this change in place, We have observed this improving the performance
significantly on Ceph read operations. As shown in
Figure~\ref{fig:rados-tcp-autotune-disabled}.

Recent versions of Ceph work around this issue by manually controlling the TCP
buffer size.

\subsubsection{Scaling on number of OSDs per server}

We also observed that using two or more client processes and many concurrent
operations are important when performing these tests.  We tested eight client
processes with 32 concurrent 4 MB objects in flight each. We created a pool
for each RADOS Bench process to ensure that object reads come from independent
pools (RADOS Bench is not smart enough to ensure that objects are not read by
multiple processes and thus possibly cached).  A sync and flush is performed
on every node before every test to ensure no data or metadata is in cache.
All tests were run with replication set to one.

In the following test, a single Ceph host drives $n$ OSDs, where $n$ increases
from one to eleven. The result is illustrated in Figure~\ref{fig:osd-scale}.
We ran the test against a single client with four concurrent processes. In this
test case, we observe that the OSD server exhibits near linear scalability up to
nine OSDs, and is still trending upwards at eleven OSDs. This suggests that we
have not reached the saturation point yet. Additional testing would require
provisioning more OSDs on the SFA10K backend.

@@

\subsubsection{Scaling on number of OSD servers}

In this test, we exercise OSD servers from one to four, driven by four hosts
each with four RADOS Bench process. Each additional OSD server adds eleven
more OSDs into play.  Figure~\ref{fig:oss-scale} summarized the results.
Comparing to Figure~\ref{fig:rados-tcp-autotune-disabled}, the read
performance improved about 10\% given eight clients instead of four, but write
performance is mostly flat for reason explained below. We also observe that
Ceph exhibits linear scaling with regard to number of servers, at least in the
given number of servers.  However, the peak performance we are seeing is about
the half of what are expecting from the SFA10K (compare to the baseline block
I/O performance number presented in Section~\ref{sec:block-io}).

For writes, the lost performance is attributed to the way Ceph performs
journaling: Ceph does not support meta-data only journaling, therefore every
write is the equivalent of a double-write: once to the data device, once to
the journaling device. This effectively cuts the observed system bandwidth in
half. That said, it does not explain the read performance -- it is a little
better than write, but still far from the theoretical maximum.

@@

Our initial tests was less than ideal as we encountered various issues.  The
full permutation of IOR parameters sweep were not explored due to I/O errors
we encountered.  We were, however, able to record results in two extreme cases
as far as transfer size is concerned: 4 KB and 4 MB, using a fixed number of
OSD servers (4) and fixed block size (8 GB), the results are illustrated in
Figure \ref{fig:ior4k} and \ref{fig:ior4m}, we make the following
observations:

@@

\subsection{Metadata Performance Evaluation}

In our particular setup, we only had one metadata server (MDS) configured.
Therefore, this is not a scalability test on the performance of Ceph clustered
MDS, which would have been very interesting.  Instead, we focus on a single MDS
performance and exercise it with up to 8 clients to observe the single MDS
performance scaling. We used
mdtest\footnote{\url{http://sourceforge.net/projects/mdtest/}} as our benchmark
tool. Mdtest parameters used for this test are:

\begin{itemize}
\item \verb!-w 1048576 -y!: for each created file, we write 1 MB data and
perform a sync operation.  This is a more realistic use case scenario than just
open, close and unlink sequence of metadata operations.

\item \verb!-n 1000!: Per client file \textit{and} directory workload. For
eight clients, the total number of files and directories in the workload is
8,000. Since we did not specify either \verb!-D! (indicates a directory only
workload) or \verb!-F! (indicates a file only workload), this is a mixture of
both directory and file workloads.

\item \verb!-d /mnt/cephfs/tmp!: we do specify a directory, but unlike under
Lustre file system, where you can have single client multiple mounts (for
increasing workload per client), here we just give the test an explicit home.

\item \verb!-u!: without this option, we are exercising shared directory; with
this option, we are exercising unique directories per test client processes.

Each test iterates five times and we are presenting the maximum out of all five iterations. 
We summarize the results as follows:

\begin{figure}[htb]
\centering
\includegraphics[width=3in]{data/mdtest-fcreate}
\caption{File creation vs.  number of clients}
\label{fig:mdtest-fcreate}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=3in]{data/mdtest-dcreate}
\caption{Directory creation vs. number of clients}
\label{fig:mdtest-dcreate}
\end{figure}

%Figure~\ref{fig:mdtest1c} is meant to give an overview on the trending of each
%directory (prefixed with d) and file (prefixed with f) operations, against the
%number of clients. Due to the scale difference, it doesn't convey the
%magnitude of lower numbers.

\item With either shared or unique directory (\verb!-u!), stat
operations for directories and files exhibit strong linear scaling. Same strong
linear scaling is also observed for file read operations.

\item While the other operations seems unaffected or performance remains the
same by the number of clients, it is not so if we zoom in, see
figures \ref{fig:mdtest-fcreate} and \ref{fig:mdtest-dcreate}:  as number of clients
increases, we observed the contention for shared directory. The performance
degradation amount to 50\% or more.

\item Though the same saturation (or degradation) trend was not observed for
file creation operation, it is likely due to lack of workload stress on MDS.

The results also show that file creation rate is significantly lower than
directory creation rate. We stipulated two contributing factors: one is the file
creation is associated with 1 MB data write and followed by a fsync operation,
which is a rather heavy weight operation compare to directory creation. Another
factor is that we obtained above results while both RADOS and file
system-level performance are not tuned for optimal, yet.

Next, we will describe the efforts and results on performance improvement at
both RADOS and file system-level. As consequence, we expect to see improved
metadata performance as well.

>>>>

%RADOS is a reliable distributed object store, the foundational component for
%CephFS file system. There are two types of scaling tests we are interested at
%the RADOS layer:
%
%\begin{itemize}
%  \item scaling on the number of OSD servers
%  \item scaling on the number of OSDs per OSD server
%\end{itemize}
%
%Our system setup poses some limitations on the scalability tests we wanted to
%perform. In particular, we currently had four OSD servers, eight clients, and
%eleven OSD servers per client. The scaling tests, therefore, will be within
%these constraints.

@@

Our early testing indicated a read performace abberation.  There were periods of high performance followed by
periods of low performance or outright stalls that could last for up to 20
seconds at a time.  RADOS bench recorded poor read performance across
different backend file systems, as shown in
Figure~\ref{fig:rados-tcp-autotune}.

We observed that outstanding operations on the clients were not being shown as
outstanding on the OSDs.  This appeared to be similar to a problem Jim Schutt
at Sandia National labs encountered with TCP autotuning in the Linux
kernel\cite{Jim:tcptune}.
TCP auto tuning enables TCP window scaling by default and automatically adjusts
the TCP receive window for each connection based on link conditions such as
bandwidth delay product.

@@

We disabled this behavior by setting 
\url{/proc/sys/net/ipv4/tcp_moderate_rcvbuf} to 0 on all nodes, improving the Ceph read performance as
shown in Figure~\ref{fig:rados-tcp-autotune-disabled}.
Recent versions of Ceph work around this issue by manually controlling the TCP
buffer size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% REMOVED for space

@@

\subsubsection{Scaling OSDs per server}

We also observed that using two or more client processes and many concurrent
operations are critical to achieve high performance.  We tested eight client
processes with 32 concurrent 4 MB objects in flight each. We created a pool
for each RADOS Bench process to ensure that object reads come from independent
pools (RADOS Bench is not smart enough to ensure that objects are not read by
multiple processes and thus possibly cached).  A sync and flush is performed
on every node before every test to ensure no data or metadata is in cache.
All tests were run with replication set to one. Figure~\ref{fig:osd-scale}
shows that an OSD server exhibits near linear scalability up to
nine OSDs, and is still trending upwards at eleven OSDs. This suggests that we
have not reached the saturation point yet. Additional testing would require
provisioning more OSDs on the SFA10K backend.
%
%In the following test, a single Ceph host drives $n$ OSDs, where $n$ increases
%from one to eleven. The result is illustrated in Figure~\ref{fig:osd-scale}.
%We ran the test against a single client with four concurrent processes. In this
%test case, we observe that the OSD server exhibits near linear scalability up to
%nine OSDs, and is still trending upwards at eleven OSDs. This suggests that we
%have not reached the saturation point yet. Additional testing would require
%provisioning more OSDs on the SFA10K backend.

@@

\subsubsection{Scaling OSD servers}

We measured performance at the client (using RADOS bench) for one to four OSD
servers.  Each additional OSD server adds eleven
more OSDs into play.  Figure~\ref{fig:oss-scale} demonstrates that read
performance improved about 10\% given eight clients instead of four, but write
performance is mostly flat.  The lost write performance compared to a single
OSD is attributed to Ceph journaling.  Ceph does not support meta-data only
journaling, therefore every write is the equivalent of a double-write: once to
the data device, once to the journaling device.  It is unclear why read
performance is so much lower than the theoretical maximum.
%This effectively cuts the observed system bandwidth in
%half. 
%That said, it does not explain the read performance -- it is a little
%better than write, but still far from the theoretical maximum.
%Comparing to Figure~\ref{fig:rados-tcp-autotune-disabled}, the read
%performance improved about 10\% given eight clients instead of four, but write
%performance is mostly flat for reason explained below. We also observe that
%Ceph exhibits linear scaling with regard to number of servers, at least in the
%given number of servers.  However, the peak performance we are seeing is about
%the half of what are expecting from the SFA10K (compare to the baseline block
%I/O performance number presented in Section~\ref{sec:block-io}).
%
%For writes, the lost performance is attributed to the way Ceph performs
%journaling: Ceph does not support meta-data only journaling, therefore every
%write is the equivalent of a double-write: once to the data device, once to
%the journaling device. This effectively cuts the observed system bandwidth in
%half. That said, it does not explain the read performance -- it is a little
%better than write, but still far from the theoretical maximum.

@@

The full permutation of IOR parameters sweep were not explored due to I/O
errors.  We were able to record results in the I/O size boundary cases:
4 KB and 4 MB, using a fixed number of
OSD servers (4) and fixed block size (8 GB), the results are illustrated in
Figure \ref{fig:ior4k} and \ref{fig:ior4m}, we make the following
observations:

@@

%% \subsection{Metadata Performance Evaluation}

%% In our particular setup, we only had one metadata server (MDS) configured.
%% Therefore, this is not a scalability test on the performance of Ceph clustered
%% MDS, which would have been very interesting.  Instead, we focus on a single MDS
%% performance and exercise it with up to 8 clients to observe the single MDS
%% performance scaling. We used
%% mdtest\footnote{\url{http://sourceforge.net/projects/mdtest/}} as our benchmark
%% tool. Mdtest parameters used for this test are:

%% \begin{itemize}
%% \item \verb!-w 1048576 -y!: for each created file, we write 1 MB data and
%% perform a sync operation.  This is a more realistic use case scenario than just
%% open, close and unlink sequence of metadata operations.

%% \item \verb!-n 1000!: Per client file \textit{and} directory workload. For
%% eight clients, the total number of files and directories in the workload is
%% 8,000. Since we did not specify either \verb!-D! (indicates a directory only
%% workload) or \verb!-F! (indicates a file only workload), this is a mixture of
%% both directory and file workloads.

%% \item \verb!-d /mnt/cephfs/tmp!: we do specify a directory, but unlike under
%% Lustre file system, where you can have single client multiple mounts (for
%% increasing workload per client), here we just give the test an explicit home.

%% \item \verb!-u!: without this option, we are exercising shared directory; with
%% this option, we are exercising unique directories per test client processes.

%% \end{itemize}

%% Each test iterates five times and we are presenting the maximum out of all five iterations. 
%% We summarize the results as follows:

%% \begin{figure}[htb]
%% \centering
%% \includegraphics[width=3in]{data/mdtest-fcreate}
%% \caption{File creation vs.  number of clients}
%% \label{fig:mdtest-fcreate}
%% \end{figure}

%% \begin{figure}[htb]
%% \centering
%% \includegraphics[width=3in]{data/mdtest-dcreate}
%% \caption{Directory creation vs. number of clients}
%% \label{fig:mdtest-dcreate}
%% \end{figure}

%% %Figure~\ref{fig:mdtest1c} is meant to give an overview on the trending of each
%% %directory (prefixed with d) and file (prefixed with f) operations, against the
%% %number of clients. Due to the scale difference, it doesn't convey the
%% %magnitude of lower numbers.

%% \begin{itemize}

%% \item With either shared or unique directory (\verb!-u!), stat
%% operations for directories and files exhibit strong linear scaling. Same strong
%% linear scaling is also observed for file read operations.

%% \item While the other operations seems unaffected or performance remains the
%% same by the number of clients, it is not so if we zoom in, see
%% figures \ref{fig:mdtest-fcreate} and \ref{fig:mdtest-dcreate}:  as number of clients
%% increases, we observed the contention for shared directory. The performance
%% degradation amount to 50\% or more.

%% \item Though the same saturation (or degradation) trend was not observed for
%% file creation operation, it is likely due to lack of workload stress on MDS.

%% \end{itemize}

%% The results also show that file creation rate is significantly lower than
%% directory creation rate. We stipulated two contributing factors: one is the file
%% creation is associated with 1 MB data write and followed by a fsync operation,
%% which is a rather heavy weight operation compare to directory creation. Another
%% factor is that we obtained above results while both RADOS and file
%% system-level performance are not tuned for optimal, yet.

%% Next, we will describe the efforts and results on performance improvement at
%% both RADOS and file system-level. As consequence, we expect to see improved
%% metadata performance as well.

### 2013-09-20 ceph-tuning.tex
fix citation to Appendix E
Brad's attempt to cut to 5 pages
So here is most of my attempt to cut this paper to 5 pages.
I amde it to about 5.7 pages.  I could go after it some more,
but this is a good first draft point to check and see if we need
to add back anything I took out.

I also added the IEEEabbrv.bib file since I couldn't build
without it.

We swept over Ceph configuration parameters to examine how different tuning
options affect performance on the DDN platform. The result of this parameter
probing is illustrated in Figure~\ref{fig:parametric}. Please refer to Appendix
E for explanations of these probed parameters.

@@

Deeper analysis with \verb!perf! showed that there was heavy lock contention during
parallel compaction in the Linux kernel memory manager.  This behavior was first
observed roughly in the kernel 3.5 time frame which was the kernel
installed on our test systems.\footnote{For more information,
please refer to \url{http://lwn.net/Articles/517082/} and
\url{https://patchwork.kernel.org/patch/1338691/}.}

We upgraded our test systems with kernel version 3.9 and performed RADOS Bench
test.  The results were extremely positive and presented in
Figure~\ref{fig:rados-kernel}. As can be seen, with the 3.9 kernel, while
there was a slight improvement on write performance, read performance improved
dramatically.  In addition to the kernel change, the amount of CephFS client
kernel read-ahead cache size was increased as well.

\begin{figure}[htb]
\centering
%\includegraphics[width=3.5in]{rados-kernel-35vs39}
\includegraphics[width=3.5in]{rados_kernel}
\caption{RADOS bench: Linux kernel version 3.5 vs. 3.9}
\label{fig:rados-kernel}
\end{figure}

@@

IOR results reflecting the read-ahead cache size change are presented in
Figure~\ref{fig:ior-kernel-39}.

@@

\subsection{Revisiting Metadata Performance}

Our previous results show that file creation rate is significantly lower than
directory creation rate. Also, we conjectured two reasons: one is the file
creation is associated with 1 MB data write and followed by a \verb!fsync!
operation.  Another reason being that we obtained above results from early
version of Ceph without all the tunings and improvement applied.

Therefore, we repeated the same test on the latest CephFS 0.64. The result is
illustrated in Figure~\ref{fig:mdtest-064-file-create}. As expected, we
observed a significant improvement on the file creation rate. We also note
that as number of clients increase, the file creation rate decreases rapidly,
which suggested some form of lock contention. This presents some interesting
issues to be investigated further.

>>>>

We swept over Ceph configuration parameters to examine how different tuning
options affect performance on the DDN platform. The result of this parameter
probing is illustrated in Figure~\ref{fig:parametric}. Please refer to
Appendix E of \cite{ceph:techreport} for explanations of these probed
parameters.

@@

Deeper analysis with \verb!perf! showed that there was heavy lock contention
during parallel compaction in the Linux kernel memory manager.  This behavior
was first observed roughly in the kernel 3.5 time frame which was the kernel
installed on our test systems.\footnote{For more information,
please refer to \url{http://lwn.net/Articles/517082/} and
\url{https://patchwork.kernel.org/patch/1338691/}.}  We upgraded our test
systems with kernel version 3.9 and performed RADOS Bench test.  The results
were dramatic:  read performance improved from 4678 MB/s to 8643 MB/s (almost
an 85\% improvement).  Write performance was also fractionally improved from
4720 MB/s to 5284 MB/s 9an almost 12\% improvement).
%extremely positive and presented in
%Figure~\ref{fig:rados-kernel}. As can be seen, with the 3.9 kernel, while
%there was a slight improvement on write performance, read performance improved
%dramatically.  In addition to the kernel change, the amount of CephFS client
%kernel read-ahead cache size was increased as well.

%\begin{figure}[htb]
%\centering
%\includegraphics[width=3.5in]{rados-kernel-35vs39}
%\includegraphics[width=3.5in]{rados_kernel}
%\caption{RADOS bench: Linux kernel version 3.5 vs. 3.9}
%\label{fig:rados-kernel}
%\end{figure}

@@

Additionally, we increased the size of the CephFS kernel client's read-ahead
chance.  IOR results reflecting the read-ahead cache size change are presented
in Figure~\ref{fig:ior-kernel-39}.

@@

\subsection{Metadata Performance Tuning}
For brevity's sake, we are omitting most of our file and directory creation
rate results.  However, over the course of our testing we did notice that in
CephFS 0.64 a significant improvement in the performance of file creation with
mdtest.  Shown in Figure~\ref{fig:mdtest-064-file-create}, we note
that as number of clients increase, the file creation rate does not scale
linearly, which suggested some form of lock contention.  Due to the large
differences in the initially observed performance, and the performance with
Ceph 0.64, opportunities for file and directory create tuning may be
substantial.
